### 2021-02-22 latex/slides/Spillover_slides.tex
Update Slides for LEPHD Brown Bag

\documentclass[aspectratio=169]{beamer}
% Metropolis Theme ------------------------------------------------------------------------------
\usetheme{metropolis} % Use metropolis theme

% Title ------------------------------------------------------------------------------
\title{Difference-in-Differences with Spatial Spillovers}
\date{\today}
\author{Kyle Butts}
% \institue{}

\input{preamble.tex}

\addbibresource{references.bib}

\usepackage{adjustbox}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{dcolumn}

% Table Highlighting
\usepackage[beamer,customcolors]{hf-tikz}
\usetikzlibrary{calc}
\usetikzlibrary{fit,shapes.misc}

% To set the hypothesis highlighting boxes red.
\tikzset{hl/.style={
    set fill color=red!80!black!40,
    set border color=red!80!black,
  },
}
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (0,1.5ex) {};%
}
\newcommand\markbottomright[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0,0) {};%
    \tikz[red, ultra thick, overlay, remember picture, inner sep=4pt]
        \node[draw, rectangle, fit=(marker-#1-a.center) (marker-#1-b.center)] {};%
}

% ------------------------------------------------------------------------------
\begin{document}

% ------------------------------------------------------------------------------
\maketitle
% ------------------------------------------------------------------------------

\begin{frame}{Spatial Spillovers}
    Researchers aim to estimate the \textbf{average treatment effect on the treated}: 
    \[
        \tau \equiv \mathbb{E} \left[ Y_{i1}(1) - Y_{i1}(0) \ \vert \ D_{i} = 1 \right]
    \]

    Estimation is complicated by \textbf{Spillover Effects}

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries such as states, counties, etc.

    \begin{itemize}
        \item e.g. large employer opening/closing in a \textbf{treated} county have positive employment effects on \textit{nearby counties}
    \end{itemize}
\end{frame}

\begin{frame}{Bias from Spatial Spillovers}

    \onslide<1->{
        The canonical difference-in-differences estimate is: 
        \only<1>{
            \[ 
                \hat{\tau} = \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right]}_{\text{Counterfactual Trend} \ + \ \tau} - 
                \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]}_{\text{Counterfactual Trend}}
            \]
        }
        \only<2>{
            \[ 
                \hat{\tau} = \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right]}_{\text{Counterfactual Trend} \ + \ \tau} - 
                \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]}_{\substack{\text{Counterfactual Trend} \\[2mm] \ + \ \text{\color{purple} Spillover on Control}}}
            \]
        }
        \only<3>{
            \[ 
                \hat{\tau} = \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right]}_{\substack{\text{Counterfactual Trend} \ + \ \tau \\[2mm] \ + \ \text{\color{red} Spillover on Treated}}} - 
                \underbrace{\hat{\mathbb{E}} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]}_{\substack{\text{Counterfactual Trend} \\[2mm] \ + \ \text{\color{purple} Spillover on Control}}}
            \]
        }
    }

    Two problems in presence of spillover effects:

    \begin{itemize}
        \onslide<2->{
            \item {\bf \color{purple} Spillover onto Control Units:}

            Nearby ``control'' units fail to estimate counterfactual trends because they are affected by treatment
        }

        \onslide<3->{
            \vspace{2.5mm}
            \item {\bf \color{red} Spillover onto other Treated Units:}

            Treated units are also affected by nearby units and therefore combines ``direct'' effects with spillover effects
        }
    \end{itemize}

\end{frame}

% Grey out overlays
\setbeamercovered{transparent}

\begin{frame}{Outline}

    \begin{itemize}
        \item[1--] Using a potential outcomes framework, I decompose the difference-in-differences estimator into three parts:
        \begin{itemize}
            \item Direct Effect of Treatment, Spillover onto Treated Units, Spillover onto Control Units
        \end{itemize}

        \pause
        \item[2--] Analyze effectiveness of common ad-hoc adjustments done by researchers
        \begin{itemize}
            \item Show that an indicator for being close to treated units remove \textit{all bias} so long as the indicator contains all units affected by spillovers
            \item `Rings' are able to estimate spillover effects while still removing all bias
        \end{itemize}

        \pause
        \item[3--] Apply framework to Urban Economics
        \begin{itemize}
            \item Revisit \citet{Kline_Moretti_2014a} analysis of the Tennessee Valley Authority

            \item Discuss how framework can reconcile conflicting findings on effect of federal Empowerment Zones \citep{Busso_Gregory_Kline_2013,Neumark_Kolko_2010}
        \end{itemize}
    \end{itemize} 
\end{frame}

\begin{frame}{Contribution}

    {\footnotesize

    \underline{Difference-in-Differences Estimation with Spillovers}

    \begin{citecolor}
        [\citet{Clarke_2017}, \citet{Berg_Streitz_2019}, and \citet{Verbitsky-Savitz_Raudenbush_2012}]
    \end{citecolor}

    \begin{itemize}
        \item I generalize the work by deriving the bias in terms of potential outcomes and show how to estimate treatment effects by removing spillover effects
    \end{itemize}

    \pause
    \underline{Potential Outcome Framework for Spillovers}

    \begin{citecolor}
        [\citet{Miguel_Kremer_2004}, \citet{Vazquez-Bare_2019}, \citet{Angelucci_DiMaro_2016}, \citet{Angrist_2014}]        
    \end{citecolor}

    \begin{itemize}
        \item These papers consider spillovers in network clusters. I focus on spatial setting
    \end{itemize}

    \pause
    \underline{Place-based Policies}

    \begin{citecolor}
        [\citet{Kline_Moretti_2014b}, \citet{Kline_Moretti_2014a}, \citet{Busso_Gregory_Kline_2013}, and \citet{Neumark_Kolko_2010}]
    \end{citecolor}

    \begin{itemize}
        \item I highlight the need to control for general equilibrium effects to properly estimate local effects of policies
    \end{itemize}
    }
\end{frame}

% ------------------------------------------------------------------------------
\section{Theory}
% ------------------------------------------------------------------------------

\begin{frame}{Potential Outcomes Framework}
    For exposition, I will label units as counties. Assume all treatment occurs at the same time (2-periods or pre-post averages).\footnote{I extend this into an event study framework in the paper, but the intuition is the same as in the $2 \times 2$ setting.}

    \begin{itemize}
        \item $Y_{it}(D_i, \textcolor{blue}{h(\vec{D}, i)})$ is the potential outcome of county $i \in \{ 1, \dots, N \}$ at time $t$ with treatment status $D_i \in \{0, 1\}$.

        \item $\vec{D} \in \{0,1\}^N$ is the vector of all units treatments.

        \pause
        \item The function $\textcolor{blue}{h(\vec{D}, i)}$ maps the entire treatment vector into an `exposure mapping' which can be a scalar or a vector.

        \pause
        \item No exposure is when $\textcolor{blue}{h(\vec{D}, i)} = \vec{0}$.
    \end{itemize}
\end{frame}

% Grey out
% \setbeamercovered{transparent}

\begin{frame}{Examples of $h_i(\vec{D})$}

    Examples of $h_i(\vec{D})$:

    \begin{itemize}
        \item \textbf{Treatment within $x$ miles:}

        $\textcolor{blue}{h(\vec{D}, i)} = max_j \ 1(d(i, j) \leq x)$ where $d(i,j)$ is the distance between counties $i$ and $j$.

        \begin{itemize}
            \item e.g. library access where $x$ is the maximum distance people will travel

            \item Spillovers are non-additive
        \end{itemize}

    \end{itemize}
\end{frame}

\imageframe{../../figures/figure-spill_within.png}

\begin{frame}{Examples of $h_i(\vec{D})$}

    Examples of $h_i(\vec{D})$:

    \begin{itemize}
        \item \textbf{Treatment within $x$ miles:}

        $\textcolor{blue}{h(\vec{D}, i)} = max_j \ 1(d(i, j) \leq x)$ where $d(i,j)$ is the distance between counties $i$ and $j$.

        \begin{itemize}
            \item e.g. library access where $x$ is the maximum distance people will travel

            \item Spillovers are non-additive
        \end{itemize}

        \vspace{2.5mm}
        \item \textbf{Number of Treated within $x$ miles:}

        $\textcolor{blue}{h(\vec{D}, i)} = \sum_{j = 1}^k 1(d(i, j) \leq x)$.

        \begin{itemize}
            \item e.g. large factories opening

            \item Agglomeration economies suggest spillovers are additive
        \end{itemize}

    \end{itemize}
\end{frame}

\imageframe{../../figures/figure-spill_within_additive.png}

\begin{frame}{Estimand of Interest}
    Estimand of Interest: \[ 
        \textcolor{green}{\tau_{\text{direct}}} \equiv \textcolor{green}{\mathbb{E}\left[ Y_{i,1}(1, 0) - Y_{i,1}(0, 0) \mid D_i = 1\right]}
    \]

    \vspace{10mm}
    This is the direct effect in the absense of exposure to spillovers.
\end{frame}

\begin{frame}{Parallel Trends}
    I assume a modified version of the parallel counterfactual trends assumption: 
    \begin{align*}
        &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 1 \big] \\
        = \ &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 0 \big],
    \end{align*}

    \vspace{5mm}
    In the \textit{complete absence of treatment }(not just the absence of individual $i$'s treatment):

    Changes in outcomes does not depend on treatment status

\end{frame}

\begin{frame}{What does Diff-in-Diff identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the diff-in-diff estimate as follows:

    \begin{align*}
        \mathbb{E}\left[\hat{\tau}\right] &= \underbrace{\mathbb{E} \left[ Y_{i1} - Y_{i0} \mid D_i = 1 \right] - \mathbb{E} \left[ Y_{i1} - Y_{i0} \mid D_i = 0 \right]}_{\text{Difference-in-Differences}} \\[3mm]
        \pause&= 
        \textcolor{green}{\mathbb{E} \left[ Y_{i1}(1, 0) - Y_{i1}(0, 0) \mid D_i = 1 \right]} \\
        &\quad + \quad 
        \textcolor{red}{\mathbb{E} \left[ Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) \mid D_i = 1 \right]} \\ 
        &\quad - \quad  
        \textcolor{purple}{\mathbb{E} \left[ Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 0 \right]} \\[3mm]
        &= \textcolor{green}{\tau_{\text{direct}}} + \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}
    \end{align*}

\end{frame}

\begin{frame}{Biased Estimate for $\tau_{\text{direct}}$}
    \[ 
        \mathbb{E}[\hat{\tau}_{\text{diff-in-diff}}] - \textcolor{green}{\tau_{\text{direct}}} = \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}    
    \]
\end{frame}

\begin{frame}{Example of Bias: Large Factory Opening}\label{slide:example_bias}
    \textbf{Control Spillover:}

    $Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) = 1$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

    \textbf{Treated Spillover:}

    $Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) = 0.5$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

    \pause
    Then, the bias term is 
    \begin{align*}
        &0.5 * \text{\% of treated counties within 40 miles of another treated county} \\
        &-1 * \text{\% of treated counties within 40 miles of a treated county}
    \end{align*}

    \hyperlink{slide:partial_identification}{\beamerbutton{Partial Identification}}
\end{frame}

% ------------------------------------------------------------------------------
\section{Estimation with Spillovers}
% ------------------------------------------------------------------------------

\begin{frame}{Monte Carlo Simulations}
    \begin{itemize}
        \item Data Generating Process 
        \begin{align*}
            y_{it} &= \mu_t + \mu_i + \textcolor{green}{\tau_{\text{direct}}} D_{it} + \textcolor{purple}{\tau_{\text{spillover, control}}} (1-D_{it}) \text{Within 40mi.}_{it} \\ 
            &\quad + \textcolor{red}{\tau_{\text{spillover, treated}}} D_{it} \text{Within 40mi.}_{it} + \varepsilon_{it}
        \end{align*}

        \item Observation $i$ is a U.S. county, year $t \in \{2000, \dots, 2019\}$, treatment $D_{it}$ turns on in 2010 and is assigned randomly or through a spatial process.

        \item $\text{Within 40mi.}_{it}$ is an indicator for having your county center of population be within 40 miles of a treated unit.

        \item $\textcolor{green}{\tau_{\text{direct}}} = 2$, $\textcolor{purple}{\tau_{\text{spillover, control}}} = 1$ and $\textcolor{red}{\tau_{\text{spillover, treat}}} \in \{0, 0.5\}$.
    \end{itemize}
\end{frame}

\begin{frame}{Source of Bias 1: Control Units}
    Simulation 1:

    \begin{itemize}
        \item The first simulation changes the probability of treatment, $D_{it}$.

        \item Assume, for now, there is only spillover on the control, i.e. $\tau_{\text{spillover,treat}} = 0$.

        \item Estimate via TWFE: \begin{equation*}
            y_{it} = \mu_t + \mu_i + \tau D_{it} + \epsilon_{it}    
        \end{equation*}

        \item The bias of our estimate is \textit{Bias} $= \hat{\tau} - 2$
    \end{itemize}

\end{frame}

\imageframe{../../figures/figure-bias_from_prob_treat_slides.png}

\begin{frame}{Solution: Removing ``contaminated'' controls}
    A common solution to the problem of spillover is to restimate on a subsample with neighboring control units removed.

    This simulation drops only units with $\text{Within 40mi.}_{it} = 1$. The exact exposure mapping is typically not observable for researchers, so this is best case scenario.
\end{frame}

\imageframe{../../figures/figure-bias_fix_slides.png}

\begin{frame}{Why Removing Controls Doesn't Work Well}
    \begin{itemize}
        \item[1.] Lose precision in treatment effect estimate by dropping control units
        \item[2.] If there are spillovers on treated units, removing controls won't remove both sources of bias. 
    \end{itemize}

    I repeat the same simulations but with $\textcolor{red}{\tau_{\text{spillover, treat}}} = 0.5$

\end{frame}

\imageframe{../../figures/figure-bias_fix_treat_slides.png}

\begin{frame}{Solution: Parametrize potential outcomes}
    Another option is to directly control for spillovers by parametrizing the exposure mapping.

    This is not observable to a researcher which raises the question:

    Are there specifications that work well even if the spillovers are misspecified?
\end{frame}

\begin{frame}{Robustness to Misspecification}
    Generate data using the same data-generating process as before but with different spillover functions:

    \[ 
        y_{\it} = \mu_t + \mu_i + 2 D_{it} + \beta_{\text{spill,control}} * (1-D_{it}) h(\vec{D}, i) + \varepsilon_{it}
    \]

    Then, I estimate each data-generating process using (potentially) misspecified $\tilde{h}(\vec{D}, i)$ and report the average estimate bias.

\end{frame}

\begin{frame}{Specifications of $h(\vec{D}, i)$}

    \begin{itemize}
        \item \textit{Within $40/80$mi.}:
        \begin{itemize}
            \item Inidicator for nearest treated unit being within $40/80$ miles.
        \end{itemize}

        \item \textit{Within $40/80$mi. (Additive)}: 
        \begin{itemize}
            \item Number of treated units being within $40/80$ miles.
        \end{itemize}

    \end{itemize}

\end{frame}

\imageframe{../../figures/figure-spill_within_large.png}
\imageframe{../../figures/figure-spill_within_large_additive.png}

\begin{frame}{Specifications of $h(\vec{D}, i)$}

    \begin{itemize}
        \item \textit{Within $40/80$mi.}:
        \begin{itemize}
            \item Inidicator for nearest treated unit being within $40/80$ miles.
        \end{itemize}

        \item \textit{Within $40/80$mi. (Additive)}: 
        \begin{itemize}
            \item Number of treated units being within $40/80$ miles.
        \end{itemize}

        \item \textit{Decay}: 
        \begin{itemize}
            \item $\max_j D_j * e^{-0.02 d(i,j)} * 1(d(i,j) < 80)$
        \end{itemize}

        \item \textit{Decay (Additive)}: 
        \begin{itemize}
            \item $\sum_j D_j * e^{-0.02 d(i,j)} * 1(d(i,j) < 80)$
        \end{itemize}

    \end{itemize}

\end{frame}

\imageframe{../../figures/figure-spill_decay.png}
\imageframe{../../figures/figure-spill_decay_additive.png}

\begin{frame}{Specifications of $h(\vec{D}, i)$}

    \begin{itemize}
        \item \textit{Within $40/80$mi.}:
        \begin{itemize}
            \item Inidicator for nearest treated unit being within $40/80$ miles.
        \end{itemize}

        \item \textit{Within $40/80$mi. (Additive)}: 
        \begin{itemize}
            \item Number of treated units being within $40/80$ miles.
        \end{itemize}

        \item \textit{Decay}: 
        \begin{itemize}
            \item $\max_j D_j * e^{-0.02 d(i,j)} * 1(d(i,j) < 80)$
        \end{itemize}

        \item \textit{Decay (Additive)}: 
        \begin{itemize}
            \item $\sum_j D_j * e^{-0.02 d(i,j)} * 1(d(i,j) < 80)$
        \end{itemize}

        \item \textit{Rings}: 
        \begin{itemize}
            \item Set of concentric rings. For each ring, indicator for nearest treated unit being within that distance bin
        \end{itemize}

        \item \textit{Rings (Additive)}: 
        \begin{itemize}
            \item Set of concentric rings. For each ring, number of treated units being within that distance bin
        \end{itemize}
    \end{itemize}

\end{frame}

\imageframe{../../figures/figure-spill_ring.png}

\begin{frame}

    \begin{table}[!tb]
        \caption{Bias from Misspecification of Spillovers}
        \label{tab:misspecification}

        \begin{adjustbox}{width = 1.1\textwidth, center}
            \begin{threeparttable}
                \begin{tabular}{@{} l rrrrrr @{}}
                    % Head
                    \toprule
                    & \multicolumn{6}{c}{Data-Generating Process} \\
                    \cmidrule{2-7}

                    & \multicolumn{1}{c}{\textbf{Within 40mi.}} & \multicolumn{1}{c}{\textbf{Within 80mi.}} & \multicolumn{1}{c}{\textbf{Within 40mi.}} & \multicolumn{1}{c}{\textbf{Within 80mi.}} & \multicolumn{1}{c}{\textbf{Decay 80mi.}} & \multicolumn{1}{c}{\textbf{Decay 80mi.}} \\
                    Specification & & & \multicolumn{1}{c}{\textbf{(Additive)}} & \multicolumn{1}{c}{\textbf{(Additive)}} & & \multicolumn{1}{c}{\textbf{(Additive)}} \\

                    % Body
                    \midrule

                    % \input{../../tables/misspecification_slides.tex}
                    TWFE (No Spillovers) & $0.258$ & $0.258$ & $0.258$ & $0.258$ & $0.258$ & $0.258$ \\ \midrule
                    \marktopleft{c1}Within 40mi. & $-0.005$ & $0.213$ & $-0.005$ & $0.176$ & $0.159$ & $0.143$ \\ 
                    Within 80mi. & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ \markbottomright{c1}
                    \onslide<2->{
                        \\ \midrule
                        Within 40mi. (Additive) & $0.043$ & $0.221$ & $-0.006$ & $0.177$ & $0.174$ & $0.143$ \\ 
                        Within 80mi. (Additive) & $0.034$ & $0.134$ & $-0.012$ & $-0.009$ & $0.099$ & $-0.010$ \\ \midrule
                        Decay 80mi. & $-0.159$ & $0.070$ & $-0.174$ & $0.014$ & $-0.009$ & $-0.033$ \\ 
                        Decay 80mi. (Additive) & $-0.023$ & $0.148$ & $-0.084$ & $0.019$ & $0.088$ & $-0.008$
                    }
                    \onslide<3->{
                        \\ \midrule
                        \marktopleft{c2}Rings (0-20, 20-30, 30-40) & $-0.005$ & $0.213$ & $-0.005$ & $0.176$ & $0.159$ & $0.143$ \\ 
                        Rings (0-20, 20-30, 30-40, 40-60, 60-80) & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ & $-0.009$ \markbottomright{c2}\\ 
                        Rings (0-20, 20-30, 30-40, 40-60, 60-80) (Additive) & $0.036$ & $0.134$ & $-0.008$ & $-0.008$ & $0.100$ & $-0.009$ 
                    }

                    \\ \bottomrule
                \end{tabular}
            \end{threeparttable}
        \end{adjustbox}
    \end{table}

\end{frame}

\begin{frame}{Spillovers as Estimand of Interest}
    Until now, we assumed our estimand of interest is $\textcolor{green}{\tau_{\text{direct}}}$.

    However, the two other spillover effects are of interest as well:
    \begin{itemize}
        \item $\textcolor{purple}{\tau_{\text{spillover, control}}}$: Do the benefits of a treated county come at a cost to neighbor counties?

        \item $\textcolor{red}{\tau_{\text{spillover, treated}}}$: Does the estimated effect change based on treatment of neighbors? 
    \end{itemize}

    To estimate the spillover effects, we have to parameterize $h(\vec{D}, i)$ function and the potential outcomes function $Y_i(D_i, h(\vec{D}, i))$.
\end{frame}

\begin{frame}{Estimation of Spillover Effects}
    To see which specifications can predict spillover effects well, I estimate the spillover effects for each control unit, $\hat{\beta}_{\text{spill, control}} * \tilde{h}(\vec{D}, i)$.

    Then calculate \[ 
        1 - \frac{
                \overbrace{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i) - \hat{\beta}_{\text{spill, control}} \tilde{h}(\vec{D}, i))^2}^{\text{Mean Square Prediction Error}}
            }{
                \underbrace{\sum_{i: D_i = 0} (\beta_{\text{spill, control}} h(\vec{D}, i))^2}_{\text{Normalization}}
            }    
    \]

    This gives the proportion of spillovers explained by $\tilde{h}(\vec{D}, i)$

\end{frame}

\begin{frame}
    \begin{table}[!tb]
        \caption{Percent of Spillovers Predicted by Specification}
        \label{tab:misspecification_mspe}

        \begin{adjustbox}{width = 1.1\textwidth, center}
            \begin{threeparttable}
                \begin{tabular}{@{} l rrrrrr @{}}
                    % Head
                    \toprule
                    & \multicolumn{6}{c}{Data-Generating Process} \\
                    \cmidrule{2-7}

                    & \multicolumn{1}{c}{\textbf{Within 40mi.}} & \multicolumn{1}{c}{\textbf{Within 80mi.}} & \multicolumn{1}{c}{\textbf{Within 40mi.}} & \multicolumn{1}{c}{\textbf{Within 80mi.}} & \multicolumn{1}{c}{\textbf{Decay 80mi.}} & \multicolumn{1}{c}{\textbf{Decay 80mi.}} \\
                    Specification & & & \multicolumn{1}{c}{\textbf{(Additive)}} & \multicolumn{1}{c}{\textbf{(Additive)}} & & \multicolumn{1}{c}{\textbf{(Additive)}} \\

                    % Body
                    \midrule

                    % \input{../../tables/misspecification_slides.tex}
                    TWFE (No Spillovers) & $0.0\%$ & $0.0\%$ & $0.0\%$ & $0.0\%$ & $0.0\%$ & $0.0\%$ \\ \midrule
                    \marktopleft{c3}Within 40mi. & $99.4\%$ & $25.9\%$ & $85.6\%$ & $38.8\%$ & $59.5\%$ & $56.1\%$ \\ 
                    Within 80mi. & $39.8\%$ & $96.2\%$ & $34.3\%$ & $71.7\%$ & $85.6\%$ & $68.0\%$ \markbottomright{c3}
                    \onslide<2->{
                    \\ \midrule
                        Within 40mi. (Additive) & $85.3\%$ & $21.2\%$ & $99.5\%$ & $40.6\%$ & $52.0\%$ & $60.7\%$ \\ 
                        Within 80mi. (Additive) & $45.8\%$ & $61.8\%$ & $47.2\%$ & $98.4\%$ & $71.0\%$ & $93.6\%$ \\ \midrule 
                        Decay 80mi. & $60.1\%$ & $82.5\%$ & $52.7\%$ & $75.8\%$ & $97.5\%$ & $82.2\%$ \\ 
                        Decay 80mi. (Additive) & $60.7\%$ & $56.9\%$ & $63.8\%$ & $93.5\%$ & $79.0\%$ & $98.7\%$ 
                    }
                    \onslide<3->{
                        \\ \midrule 
                        \marktopleft{c4}Rings (0-20, 20-30, 30-40) & $98.4\%$ & $23.7\%$ & $85.9\%$ & $37.5\%$ & $58.9\%$ & $56.2\%$ \\ 
                        Rings (0-20, 20-30, 30-40, 40-60, 60-80) & $96.6\%$ & $91.7\%$ & $84.2\%$ & $72.7\%$ & $91.9\%$ & $78.4\%$ \\ 
                        Rings (0-20, 20-30, 30-40, 40-60, 60-80) (Additive) & $83.5\%$ & $57.4\%$ & $97.6\%$ & $95.0\%$ & $73.5\%$ & $94.9\%$\markbottomright{c4} 
                    }

                    \\ \bottomrule
                \end{tabular}
            \end{threeparttable}
        \end{adjustbox}
    \end{table}

    \onslide<3->{
        \textbf{Donuts perform best at estimating spillover effects.}

        \textbf{It is important to get Additive vs. Non-Additive correct.}
    }
\end{frame}

% ------------------------------------------------------------------------------
\section{Application in Urban Economics}
% ------------------------------------------------------------------------------

\begin{frame}{Tennessee Valley Authority}
    \citet{Kline_Moretti_2014a} look at the long-run impacts of the Tennessee Valley Authority (TVA).

    \begin{itemize}
        \item The TVA was a large-scale federal investment started in 1934 that focused on improving manufacturing economy.

        \item The program focused on electrification through dams and transportation canals in order to improve the manufacturing economy

        \item Hundreds of dollars spent anually per person
    \end{itemize}
\end{frame}

\begin{frame}{Identification}
    \citet{Kline_Moretti_2014a} run the county-level difference-in-differences  specification: 
    \begin{equation}\label{eq:tva_spillover}
        y_{c, 2000} - y_{c, 1940} = \alpha + \text{TVA}_c \tau + X_{c, 1940} \beta + (\varepsilon_{c, 2000} - \varepsilon_{c, 1940})
    \end{equation}

    \begin{itemize}
        \item $y$ are outcomes for agricultural employment, manufacturing employment, and median family income.
        \item $\text{TVA}_c$ is the treatment variable
        \item $X_{c, 1940}$ allow for different long-term trends based on covariates in 1940. 
    \end{itemize}

    They trim the sample using a logit regression to predict treatment using $X_{c,1940}$ and then keep control units in the top 75\% of predicted probability.
\end{frame}

\begin{frame}{Spillovers in the TVA Context}
    In our context, there is reason to believe spillovers can occur to nearby counties

    \begin{itemize}
        \item Agricultural employees might be drawn to hire wages for new manufacturing jobs in Tennessee Valley (negative spillover on control units)

        \item Manufacturing jobs that would have been created in the control units in the absence of treatment might move to the Tennessee Valley (negative spillover on control units)

        \item Since manufacutring jobs pay higher than agricultural jobs, the above two spillover effects can cause changes in median family income
    \end{itemize}
\end{frame}

\begin{frame}{Specification including spillovers}
    \begin{equation}
        \Delta y_c = \alpha + \text{TVA}_i \tau + \sum_{d \in \text{Dist}} \text{Ring}(d)\delta_d + X_{i, 1940} \beta + \Delta \varepsilon_c
    \end{equation}

    \begin{itemize}
        \item $\text{Ring}(d)$ is a set of indicators for being in the following distance bins (in miles) from the Tennessee Valley Authority: 
        \[ d \in \left\{ (0, 50], (50, 100], (100, 150] \right\} \]
    \end{itemize}
\end{frame}

\imageframe{../../figures/figure-tva-sample_slides.pdf}

\begin{frame}
\begin{table}[ht]
    \caption{Effects of Tennessee Valley Authority on Decadel Growth}
    \label{tab:tva}
    \renewcommand{\arraystretch}{1.2}

    \begin{adjustbox}{width = 1\textwidth, center}
        \begin{threeparttable}
            \begin{tabular}{@{} lc@{\extracolsep{20pt}}c@{\extracolsep{4pt}}ccc @{}}
                % Head
                \toprule

                & \multicolumn{1}{c}{\textbf{Diff-in-Diff}} & \multicolumn{4}{c}{\textbf{Diff-in-Diff with Spillovers}} \\ 
                \cmidrule{2-2} \cmidrule{3-6} 
                & & & TVA between & TVA between & TVA between \\ 
                \textit{Dependent Var.} & TVA & TVA & 0-50 mi. & 50-100 mi. & 100-150 mi. \\

                % Body
                \midrule

                Agricultural employment     & $-0.0514^{***}$& $-0.0678^{***}$& $-0.0310^{**}$ &    $-0.0112$   & $-0.0252^{***}$\\
                &   $(0.0087)$   &   $(0.0102)$   &   $(0.0123)$   &   $(0.0094)$   &   $(0.0084)$

                \onslide<2->{
                    \\
                    Manufacturing employment    & $0.0560^{***}$ &  $0.0461^{**}$ &    $-0.0104$   &    $-0.0128$   &  $-0.0248^{*}$ \\
                    &   $(0.0187)$   &   $(0.0210)$   &   $(0.0205)$   &   $(0.0257)$   &   $(0.0147)$
                }
                \onslide<3->{
                    \\
                    Median family income        & $0.0191^{***}$ &  $0.0196^{**}$ &    $0.0056$    &    $-0.0062$   &    $-0.0011$   \\
                    &   $(0.0067)$   &   $(0.0077)$   &   $(0.0079)$   &   $(0.0060)$   &   $(0.0032)$ 
                }

                \\ \bottomrule
            \end{tabular}

            % Notes 
            \begin{tablenotes}\footnotesize

                \item $^{*} p< 0.1$; $^{**} p < 0.05$; $^{***} p < 0.01$.
            \end{tablenotes}
        \end{threeparttable}
    \end{adjustbox}
\end{table}

\end{frame}

\begin{frame}{Identification Strategies and Place-Based Policies}

    The literature on federal Enterprise Zones, place-based policy that gives tax breaks to businesses that locate within the boundary, has found conflicting results, suggesting positive or near-zero effects of the program \citep{Neumark_Young_2019}.

    \begin{itemize}
        \pause
        \item \citet{Busso_Gregory_Kline_2013} compare census tracts in Empowerment Zones to census tracts that qualified and were rejected from the program. They find significant large reduction of poverty.

        \pause
        \item \citet{Neumark_Kolko_2010} compare census tracts in Empowerment Zones to census tracts within 1,000 feet of the Zone. They find near-zero effects on poverty.
    \end{itemize}

    \pause
    My framework can explain both of these results. If census tracts just outside the Empowerment Zones also benefit from the policy, then the estimates of \citet{Neumark_Kolko_2010} are attenuated towards zero

\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}
    \begin{itemize}
        \item I decomposed the TWFE estimate into the direct effect and two spillover terms

        \item I showed that a set of concentric rings allows for estimation of the direct effect of treatment and they are able to model spillovers well

        \item For place-based policies, I show the importance of considering spatial spillovers when estimating treatment effects

        \item More generally, identification strategies that use very close control units in order to minimize differences in unobservables should consider the problems with treatment effect spillovers.
    \end{itemize}
\end{frame}

% ------------------------------------------------------------------------------
\appendix\section{Appendix}
% ------------------------------------------------------------------------------

\begin{frame}{Partial Identification}\label{slide:partial_identification}
    Researchers can use the above method to create bounds for the biases resulting from spillovers. The choices should follow from the empirical context at hand.

    To formalize this, I follow the literature on partial identification for treatment effects
    \begin{citecolor}
        [\citet{Armstrong_Kolesár_2020}, \citet{Rambachan_Roth_2020}, and \citet{Manski_Pepper_2018}]
    \end{citecolor}
\end{frame}

\begin{frame}{Partially-identified Set}
    Following \citet{Rambachan_Roth_2020} and \citet{Manski_Pepper_2018}, researchers can provide bounds to the maximum spillovers.

    Let $\Delta$ denote the interval of potential biases: 
    \begin{align*}
        \Delta &\equiv [\underline{\Delta}, \overline{\Delta}] \\
        &= \left[ \min \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}, \ \max \tau_{\text{spill,treated}} - \tau_{\text{spill,control}} \right] 
    \end{align*}

    The partially identified set is formed by \[ 
        \left[\hat{\tau} - \underline{\Delta} , \ \hat{\tau} + \overline{\Delta} \right]    
    \]

\end{frame}

\begin{frame}{Bias-adjusted Inference}

    Inference can be done following \citet{Rambachan_Roth_2020} by creating a Fixed-Length Confidence Interval: 
    \[ 
        \left[
            \hat{\tau} - \underline{\Delta} - z_{1 - \alpha/2} \text{SE}_{\hat{\tau}}, \ \hat{\tau} + \overline{\Delta} + z_{1 - \alpha/2} \text{SE}_{\hat{\tau}}
        \right]
    \]

    These standard errors should account for the spatial structure of outcomes following \citet{Conley_1999}.

    \hyperlink{slide:example_bias}{\beamerbutton{Back to Example of Bias}}
\end{frame}

\end{document}